Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-01-26 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

Rotational motion about a single axis is easiest to account for.  
When the power spectra of elliptical polystyrene particles was 
fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
assumed that the rotational motion was purely in the transverse
plane. As such they did not have to account for any variance in 
the trapping strength due to orientation nor did they need to
consider non-periodic rotational behaviour. In the case where 
rotational motion is stochastic the problem is more complex. 
For example, when an optical fibre trap characterisation 
technique was implemented by Saffron \textit{et al} 
\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
light scattering to characterise both the axial and lateral 
trap stiffness acting on microspheres. The only drawback 
admitted to in their work was that the technique was constrained 
to isotropic scatters as their theoretical model for describing 
the auto-correlation function was predicated on the fact that 
any variations in the signal are due to the particles 
translational motion within the confines of a cylindrical trap 
\cite{BarZiv1997}. Where the upper limit of the cylindrical 
trap is given by the Rayleigh range ($z_R = n\omega_0/NA$).
However, as demonstrated by the results from Chapter 4.1, 
the axial traps of spherical aggregates is often situated 
far beyond the Rayleigh range (for a 1.2 NA laser this is 
$\pm5.985 \mu m$).

To accurately describe the full dynamics of an anisotropic
scatterer its clear that additional characterisation 
techniques are necessary. In this chapter we demonstrate 
two such techniques, quadrant photo-diode simulation, and 
an optical fibre static light scattering arrangement. The 
former is based on the work of Rohrbach and Stelzer 
\cite{Rohrbach2002}, where their work is limited to single 
spheres, we extend the possible particle space to allow for 
any theoretical combination of spheres. The results from
chapter 4 demonstrate that the rotational motion causes 
significant deviations between the computed trapping strength
and the expected trapping strength according to Lorenz-Mie
theory. By utilizing vector regression we can accurately 
recreate a particle's trajectory simply from the QPD signal, 
both in the transverse plane and along the beam axis.

The latter method builds upon \cite{BarZiv1997} but now 
accounts for the full rotational motion. This was initially
developed in tandem with the work conducted in chapter 3,
with the idea being that if a crystal nucleus formed on a 
rotating sphere how would that nucleus influence the motion
of the rotating sphere. Using static light scattering we can 
map the outputted signal from a optically trapped dimer to 
its expected orientation in real time. Thus allowing us to 
estimate the optical torque being applied to the dimer by 
the trapping beam. As demonstrated in chapter 4, the 
influence of a circularly polarised beam can result in 
gyroscopic precession for specifically sized dimers.

We start by outlining the working principle behind the
optical fibre detection system, before demonstrating 
how it can be used to instantaneously predict 
orientational information on a trapped particle. We then
consider common factors of error in the characterisation
process, such as signal error, and incorrect particle 
sizing. Both of which have a significant impact on the 
performance of the model, but by utilising Bayesian 
inference, and time averaging we can better refine our 
model.

@@

\subsection{Dimer}
The dimer is defined by two spheres with refractive indices
of 1.59 and suspended in water ($n_{med} = 1.33$). \textit{MSTM}
requires a target origin for setting the scattering expansion, as
with our simulations we set it as the dimer's centre of diffusion.

@@

While \textit{mstm} does provide the amplitude matrix it is not
scaled to account for distance from the target to the detector.
We amend this by simply scaling according to the inverse square 
law. As the detector is moved further from the target the 
intensity distribution across the entire detector will decrease
substantially. Because of the fall off it is more computationally 7
efficient to instead consider only the central pixel of the 
detector rather than across the entire surface. If we consider 
the difference between the average intensity and the intensity 
of the central point we can see that after not even $20 \mu$ 
away from the target particle the average intensity across the 
surface is within $0.1\%$ of the central pixel. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{fall_off_with_distance.png}
	\caption{Normalised difference between the central pixel intensity
	$I(\theta, \phi)$ and the average pixel intensity over the surface
	of the optical fibre $\langle I \rangle$ as a function of distance.
	4 different optical detectors were tested (spherical coordinates 
	are given in the legend). While the choice of coordinates is 
	arbitrary they all demonstrate the same exponential decay. Dashed
	lines indicate the difference after $20\mu m$ from the trap focus,
	in most practical cases getting an optical fibre any closer would
	disrupt the trap stability.}
\end{figure}

By moving the detectors further back we can treat each detector 
as single point and compute the intensity at the centre of the 
detector rather than consider the full surface. For computing 
the instantaneous orientation we only consider the intensity 
at the centre of the detector.

\section{Interpretation of scattering data into orientation estimates}
We now discuss the methodology for computing the instantaneous 
orientation of an optically bound particle based on scattering
detected by a finite grouping of optical fibres.

@@

Our goal is to determine the orientation of the trapped dimer 
based on the measured intensity $I(\hat{n}, \ \theta_k,\ \phi_k)$. 
Rather than aim an exact estimate of the dimer's orientation, 
for the purposes of interpretation of the scattering and 
optimisation of the measurement setup it is more convenient 
to discretize the possible orientation space into a number 
of possible reference orientations. We can then use each as 
'classification categories' in a neural network methodology 
to map scattering data to orientation (see below for further 
discussion). Here we choose $\textit{n}_{ref} \ = \ 30$ 
reference orientations $\hat{\bf n}_{\alpha}$  evenly 
distributed on a unit sphere \cite{Reyuthor2006} 
(Figure~\ref{fig:dimer}b) leading to a maximum nearest 
neighbour spacing between two neighbouring reference 
orientations of 0.895 radians. Using \textit{mstm} we 
compute the raw intensities at each of the measurement angles 
that would be generated by a dimer in each reference 
orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the 
number and position of detection fibres is technically arbitrary 
there are several constraining factors that limit our ability 
to infer useful information from the trapped object, see Section~\ref{sec:brownian} for a detailed breakdown of our 
choice of detection angles. In order for a neural network
to properly weight each signal the raw intensities are 
normalized according to:
\begin{align}
	\label{eq:scale}
	y_k(\hat{\bf n}_\alpha)
	= 
	\frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
	{\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}

@@

Fig~\ref{fig:mixing} shows two clusters of orientation 
vectors and their respective signals collected from 3 detectors
- the points have been coloured based on their proximity 
to the centre of their respective cluster. While the 
orientation space appears tightly packed and ordered the 
signal space quickly spreads out in an asymmetric fashion. 
Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
mapping can intersect itself which only further increases 
the complexity. While in some instances the mapping between 
one reference orientation and another is discrete, in other 
instances the mapping becomes far more complex to discern.
\newpage 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.8\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3a.png}
	\end{subfigure}
	\begin{subfigure}{0.8\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3b.png}
	\end{subfigure}
	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals. Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
	\label{fig:mixing}
\end{figure}
\newpage
Nevertheless, at least where the uncertainty in signal measurements 
is low, we can predict the orientation from the scattering scattering 
by utilising computational techniques such as neural networks. We 
thus utilised the Python machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a neural network for identifying the 
dimer's orientation from its light scattering signal. The network 
was trained by generating a database of random orientation vectors, 
calculating the corresponding light scattering signals, and then 
using the network to estimate the probability of a given signal coming 
from a dimer in a given reference orientation. The network's loss 
function was evaluated and used to improve the estimation, the 
network being trained until the improvement in the loss function 
was less than 0.0001.

@@

The relative entropy is a measure if how spread out our distribution
is. With larger values of $D_{KL}$ indicating that our prediction is 
spread out so the correct orientation is unsure about $\hat{\bf n}_
{best}$. Whereas a low value of $D_{KL}$ means that our model is more 
confident in $\hat{\bf n}_{best}$. The only draw back being that the 
relative entropy does nothing to indicate accuracy, only how closely 
our distribution matches our ideal result. This will be better 
exemplified later in \ref{sec:test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Brownian Simulation}
\label{sec:brownian}

@@

When all of the detectors lie in the same plane the expected signal 
can appear identical despite the dimer being in completely different orientations. It should be noted that these pairs are reflected in 
one or more axis which suggests that these are due to the arrangement 
of our detectors. More specifically, if the detectors are placed say 
in the x-y plane then only when the dimer is pointed nearly fully 
upright will the expected signal be entirely unique. This is not an 
issue with typical analytical static light scattering experiments 
because the signal being collected is averaged over a population of 
particles all in different orientations. Whereas in our set up we are concerned with the scattering of a single particle whose scattering is orientationally dependant.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing the Neural Network}
\label{sec:test}
We applied Eq.~\eqref{eq:bayes}, taking the reference orientation 
with the highest probability  as our estimate of the dimer's 
instantaneous orientation $\hat{\bf n}_{est}$. To visualise the 
model's performance we plotted the orientation components of our 
estimation $\hat{\bf n}_{\alpha}$ and the dimer's \emph{actual} 
instantaneous orientation $\hat{\bf s}$ versus time. For comparison, 
we also plotted the orientation components of the closest reference orientation, $\hat{\bf n}_{best}$. Assuming a uniform prior of the 
reference orientations $p(\hat{\bf n}_\alpha)$ the neural network's predictions ($\hat{\bf n}_{\alpha}$ from Eq.~\eqref{eq:bayes}) are 
at times reasonable, but there are significant large and random jumps 
away from the correct result (Fig.~\ref{fig:uniform}). 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{fig5.png}
	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity. Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the blue line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$. Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
	\label{fig:uniform}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Accounting for sources of error in light scattering measurements}
When it comes to analysing light scattering from any size particle,
error analysis becomes a significant factor. Typically this can be 
accounted for by averaging over long periods of time to get an 
assessment of the steady state conditions of the target particle. 
However in our case where we wish to know the instantaneous 
orientation, we instead have to rely on our understanding of how 
uncertainty can effect our model's performance. We identified two 
areas which are likely sources of error in our estimation: firstly, 
an incorrect modelling of the target particle, and secondly, signal 
noise arising from experimental factors. We highlight how we address 
these areas below. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subparagraph{Impact of incorrect dimer sizing}
\label{sec:lam}

@@

Notably the increasing the number of reference orientations had a greater effect when our neural network was trained on a 1:1.95 dimer than a 1:2.05 dimer. This suggests that overshooting our size estimate will be less detrimental to our estimation. Notably if the our sizing is off the neural network does not predict a smooth motion within the trap; instead predicting that the dimer is jumping back and forth between different orientations. This suggest that we can narrow down our estimate of the particle's size by assessing how the dimer is reorienting within the trap, as we should expect a smooth continuos prediction. Since we are working with a spherical dimer it also stands to reason that techniques such as image analysis could be used in part to address this, so long as the trapped entity is sufficiently illuminated. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subparagraph{Impact of measurement noise on model predictions}
\label{sec:epsilon}

@@

>>>>

In chapter 4 we showed that while machine learning can be used 
to accurately predict the translational displacements of an 
aggregate particle, it fails when trying to account for the 
rotational behaviour of said particle. In this chapter we 
build upon \cite{BarZiv1997} but now account for the full 
rotational motion. This was initially developed in tandem with 
the work conducted in chapter 3, with the idea being that if a 
crystal nucleus formed on a rotating sphere how would that 
nucleus influence the motion of the rotating sphere. Using 
static light scattering we can map the outputted signal from a 
optically trapped dimer to its expected orientation in real 
time. Thus allowing us to estimate the optical torque being 
applied to the dimer by the trapping beam. As demonstrated 
in chapter 4, the influence of a circularly polarised beam 
can result in gyroscopic precession for specifically sized 
dimers.

We start by describing a quadrant photo diode system 
simulated the power spectrum derived from a particle 
trajectory. We demonstrate that even in idealised 
circumstances the rotational motion is not revealed 
from the collected power spectrum. We then go about 
defining an optical fibre detection system from a 
mathematical perspective. Before demonstrating how it 
can be used to instantaneously predict orientational 
information on a trapped particle. We then consider 
common factors of error in the characterisation process, 
such as signal error, and incorrect particle sizing. 
Both of which have a significant impact on the performance 
of the model, but by utilising Bayesian inference, and 
time averaging we can better refine our model.

\section{Characterisation of asymmetric dimer dynamics via PSD analysis}
As discussed in \ref{sec:simulated_QPD}, one of the methods 
developed to work in conjunction with \cite{Vigilante2020} 
is a simulated quadrant photo diode for as a position 
detection system. While it is possible to extract all of 
the relevant dynamical information from a simulation, 
confirming the same behaviour in an experimental setting 
can be challenging if dealing with a non-birefringent 
anisotropic scatterer. The QPD is composed of 4 photodiodes 
that measure the intensity of light incident on their 
surfaces. Using \textit{ott} we can define a region in the 
far-field that is analogues to the surface of the QPD surface. 
Evaluating the electric field across this surface gives us 
an approximation of the QPD signal outputted in an experimental 
situation. This does not provide a one-to-one result however 
as hardware errors (i.e. internal resistance, external light 
sources, and vibrations) would distort the signal somewhat. We 
can use this as a means to examine the limitations of using 
back-focal plane interferometry as a means for characterising 
anisotropic scatterers.

As a benchmark we start by considering a single sphere within 
an optical trap. A single polystyrene sphere suspended in water 
($a=1\mu m$, $n_p=1.59$, $n_m=1.33$) was trapped by a focused 
Gaussian beam ($NA=1.25$) using circularly polarised light. For 
the sake of time efficiency the trajectory was sampled every 10
time steps, meaning the upper bound on the power spectra is
$f_{Nyq}=f_{sample}/2=5000\ Hz$. To optimise the frequency window
we fitted the power spectra using the aliased Lorentzian 
(Eq.~\eqref{eq:aliased_lorentzian}). 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{PSD_sphere.png}
	\caption{Recorded power spectra fitted to eq.~
		\ref{eq:alaised_lorentzian}, scattered points represents 
		the blocked data ($n_b=100$). Corner frequency for the 
		Lorentzian curves are reported in the legend.}
	\label{fig:psd_sphere}
\end{figure}

As shown in fig.~\ref{fig:psd_sphere}, the two power spectra 
report different corner frequencies which would be indicate 
that the trap is not perfectly circular. We can use both 
\textit{ott} and the trajectory itself to derive an estimation 
of the trap geometry. Additionally, we fitted the 
same Lorentzian to the sphere's positional data, this would be a
situation when the QPD signals are completely uncorrelated with 
one another and there is a constant ratio between the QPD signal 
and the sphere's position. The corner frequencies and corresponding 
trap stiffness are reported below:

\begin{center}
	\captionof{table}{QPD fitting for single sphere}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimates} 
		& \multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory 
			fitting}\\
		\hline
		$f_c\ [Hz]$ & 447 & 450 & \parbox{1cm}{\centering 439} & 474 
		& \parbox{1.25cm}{\centering 523} & 513 \\
		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.30 & 9.65 & 8.67 & 8.61 \\
		$\kappa\ [pN/\mu m]$ & 53.05 & 53.40 & 51.96 & 56.09 & 61.94 & 60.7 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{8.16 \%} &
		\multicolumn{2}{|c|}{27.17 \%} &
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline

	\end{tabular}
\end{center}

Where $\sigma(f_c)$ is computed from \cite{BergSoerensen2004}, 
where the variance is based upon our choice of frequency 
boundaries ({$f_{min}:f_{max}$} = {$100\ Hz: 5000\ Hz$}). And 
the ellipticity of the beam is given by $e = (1-\kappa_y/
\kappa_x)^{0.5}$ and is a measure of the symmetry of the beam 
wavefront. Its clear from these initial results that the QPD 
is more sensitive to changes along the y-axis than the x-axis 
when compared to the direct \textit{ott} calculations. This is 
somewhat reflected in the trajectory results. Typically, even 
an industrial Gaussian beam will produce an elliptical diffraction 
limit spot when heavily focused; in their tutorial for optimizing 
the PSD analysis, Berg and Sorensen reported a ellipticity of 
around 15 \% after a total calibration time of 80 seconds 
\cite{BergSoerensen2004}.

The reason for the discrepancies between all 3 methods is due 
to what is actually being measured. The \textit{ott} estimates 
are simply looking at the differences in the trapping strength 
along the Cartesian axis'. If calibrated over an long enough 
time frame you would expect that the resulting power spectra 
would exactly mirror the \textit{ott} predictions. However over 
a short calibration time a QPD will only sample a small trajectory 
that does not fully explore the trapping plane. There is a clear 
trade off in terms of accuracy and computation time as shorter 
calibration runs are computationally more efficient but prone to 
errors.

With this in mind, let us consider a symmetric dimer that is 
optically trapped by the same Gaussian beam. Not only does 
the dimer's equilibrium position change but it is subjected 
to rotational motion due to its unequal moments of inertia. 
This is reflected in the calibration results using the simulated 
QPD, where we see a drastically different estimation between the 
\textit{ott} estimate and the QPD estimate.

\begin{center}
	\captionof{table}{QPD fitting for symmetric dimer}
	\label{tab:dimer}
	\begin{tabular}{ |c|c|c|c|c|c|c| } 
		\hline
		Fitting parameter & \multicolumn{2}{|c|}{\textit{ott} estimate} & 
		\multicolumn{2}{|c|}{QPD fitting} & \multicolumn{2}{|c|}{Trajectory fitting} \\
		\hline
		$f_c\ [Hz]$ & 445 & 409 & \parbox{1cm}{\centering 431} & 424 
		& \parbox{1.25cm}{\centering 274} & 285 \\
		$\sigma(f_c)\ [Hz]$ & --- & --- & 9.22 & 9.16 & 7.82 & 7.91  \\
		$\kappa\ [pN/\mu m]$ & 52.82 & 48.54 & 51.13 & 50.26 & 32.45 & 33.75 \\
		\hline
		Ellipticity &
		\multicolumn{2}{|c|}{28.5 \%} &
		\multicolumn{2}{|c|}{12.7 \%} & 
		\multicolumn{2}{|c|}{13.8 \%} \\
		\hline
	\end{tabular}
\end{center}

Now we see that the \textit{ott} predicts a more elliptical 
trap compared to the QPD model which says the trap is far
more symmetrical while trapping a symmetric dimer. A potential 
reason that \textit{ott} no longer expects a circular trap 
is because that unlike a sphere, the force displacement 
curve is not strictly harmonic. If we consider a dimer that 
is some distance from the beam axis, the sphere closest to 
the trap focus will experience a slightly greater force 
compared to the sphere further from the trap. As such its not
accurate to assume that the external force $F(x) = \kappa x$
instead we must consider that it is a function of both the 
position and orientation simultaneously.

This coupling of the translational and orientational motion 
has been highlighted previously \cite{Vigilante2020}, but 
there effects have not been demonstrated in the context of
an experimental situation. We can see from the difference in
trajectory and QPD fittings that while our description of the
trap shape is very similar (both methods give similar ellipticity
values) but the magnitude of the trap strength are significantly
different. This is partially explained by the fact that the
QPD is not actually measuring the position of the particle
but instead the intensity distribution of the total field 
incident on the QPD surface. In which case if the dimer is 
rotated slightly the QPD signal will change even if the dimers
centre of diffusion remains stationary.

\subsection{QPD for angular displacement detection}
The next logical step is to consider whether or not a QPD can at 
all be utilised for detection of rotational motion. This has been
shown for nano-particles that exhibit periodic rotational motion
using by considering the difference in diagonal quadrants. The 
benefit of detecting rotational motion is that one can begin to 
build a better understanding about how angular momentum is 
transferred to a dimer which could be extended to more complex
particles.

At just a cursory glance it would appear that there is a clear
relationship between the QPD signal and the orientation of the 
dimer. If we look at the QPD signal produced by a dimer as it is 
rotated we can see that it is equally sensitive to rotations in
either the X-Z or Y-Z plane but there is no change in the signal 
when rotated by in the X-Y plane.

However if we compare this to fig.~\ref{fig:rohrbach} we can see 
that the the signal change is dwarfed by the translational motion. 
So while their is a clear relationship between the two trying to
discern between translational and rotational contributions to the
QPD signal is not a simple task. This is significant because it 
means that with regards to the power spectrum fittings in table 
\ref{tab:dimer} there is no way of ensuring that the measured 
trap strengths are accurate. The variance $\sigma(f_c)$ is only 
the variance based upon our fitting parameters and not on the 
actual dynamics of the dimer. In the case where the particle's
shape is not known prior its not possible to determine whether 
any inaccuracies in the power spectra are due to the fit or due
to the particle's motion.

The dependence of $S_x$ and $S_y$ and the dimer's orientation
(defined by the spherical angles $\theta$, $\phi$) is therefore
not clear. To that end, we utilised machine vector regression,
which takes as an input the voltage from the 4 quadrants and 
tries to fit that to the dimer's trajectory. Fitting the QPD
signal to the dimer's position shows promising results, shown in
fig.~X.X is a prediction of a symmetric dimer's position based
on the QPD signal. With the actual displacement on the x-axis 
and the predicted result y-axis, and the dashed line represents 
the y=x line - indicating an ideal prediction.

Fig~X.X shows that its relatively trivial to predict a particle's
position based off the QPD signal, even in axial direction which
has not been done previously. It should be noted that these 
displacements are relative to the trap focus and so accurate 
tracking of the beam movements needs to be taken. This result is
partially due to the fact that displacement and signal units are
linearly related by \eqref{eq:conversion_factor}. Interestingly 
when we run the same protocol for different sized dimer's there 
seems to be a cut off point when the fit begins to fail. Plotting
the $R^2$ for each fit vs its size ratio shows that beyond a size
ratio of 5 the machine learning begins to fall off significantly.

This is surprising because we would expect that as the second 
sphere shrinks the dynamics should more closely approximate that 
of a single sphere. We can compare our

@@

\subsection{Dimer}
The dimer is defined by two spheres with refractive indices
of 1.59 and suspended in water ($n_{med} = 1.33$). \textit{MSTM}
requires a target origin for setting the scattering expansion, 
there is no agreed upon solution for where the target origin 
should be relative to the sphere positions. In chapter 4 we showed
that altering the target origin would drastically alter the 
tmatrix produced

@@

While \textit{mstm} does provide the amplitude matrix it is not
scaled to account for distance from the target to the detector.
We amend this by simply scaling according to the inverse square 
law. As the detector is moved further from the target the 
intensity distribution across the entire detector will decrease
substantially. Because of the fall off it is more computationally
efficient to instead consider only the central pixel of the 
detector rather than across the entire surface. If we consider 
the difference between the average intensity and the intensity 
of the central point we can see that after not even $20 \mu$ 
away from the target particle the average intensity across the 
surface is within $0.1\%$ of the central pixel. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{fall_off_with_distance.png}
	\caption{Normalised difference between the central pixel intensity
	$I(\theta, \phi)$ and the average pixel intensity over the surface
	of the optical fibre $\langle I \rangle$ as a function of distance.
	4 different optical detectors were tested (spherical coordinates 
	are given in the legend). While the choice of coordinates is 
	arbitrary they all demonstrate the same exponential decay. Dashed
	lines indicate the difference after $20\mu m$ from the trap focus,
	in most practical cases getting an optical fibre any closer would
	disrupt the trap stability.}
\end{figure}

Given the trapping beam's beam waist is approximately $0.54 \mu m$,
a optical fibre would need to be a minimum of $150 \mu m$ from the 
trap centre in order to prevent noise from the trapping laser itself.
At such a distance the variation in the signal is so small that it is 
easier to evaluate the intensity at the centre of the detector instead
of across the entire surface. This is how the instantaneous orientation
is predicted later on in the chapter.

\section{Interpretation of scattering data into orientation estimates}
We now discuss the methodology for computing the instantaneous 
orientation of an optically trapped particle based on scattering
detected by a finite grouping of optical fibres.

@@

Our goal is to determine the orientation of the trapped dimer 
based on the measured intensity $I(\hat{n}, \ \theta_k,\ \phi_k)$. 
Rather than aim an exact estimate of the dimer's orientation, 
for the purposes of interpretation of the scattering and 
optimisation of the measurement setup it is more convenient 
to discretize the possible orientation space into a number 
of possible reference orientations. We can then use each as 
'classification categories' in a neural network methodology 
to map scattering data to orientation (see below for further 
discussion). Here we choose $\textit{n}_{ref} \ = \ 30$ 
reference orientations $\hat{\bf n}_{\alpha}$  evenly 
distributed on a unit sphere \cite{Reyuthor2006} 
(Figure~\ref{fig:dimer}b) leading to a maximum nearest 
neighbour spacing between two neighbouring reference 
orientations of 0.895 radians.

Using \textit{mstm} we compute the raw intensities (see 
\eqref{eq:intensity}) at each of the measurement angles 
that would be generated by a dimer in each reference 
orientation, $I_s(\hat{\bf n}_{\alpha}, \theta_k)$. While 
the number and position of detection fibres is technically 
arbitrary there are several constraining factors that limit 
our ability to infer useful information from the trapped 
object, see Section~\ref{sec:brownian} for a detailed 
breakdown of our choice of detection angles. In order for 
a neural network to properly weight each signal the raw 
intensities are normalized according to:
\begin{align}
	\label{eq:scale}
	y_k(\hat{\bf n}_\alpha)
	= 
	\frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
	{\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}

@@

Fig~\ref{fig:mixing} shows two clusters of orientation 
vectors and their respective signals collected from 3 detectors
- the points have been coloured based on their proximity 
to the centre of their respective cluster. While the 
orientation space appears tightly packed and ordered the 
signal space quickly spreads out in an asymmetric fashion. 
Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
mapping can intersect itself which only further increases 
the complexity. While in some instances the mapping between 
one reference orientation and another is discrete, in other 
instances the mapping becomes far more complex to discern.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.74\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3a.png}
	\end{subfigure}
	\begin{subfigure}{0.74\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3b.png}
	\end{subfigure}
	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals. Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
	\label{fig:mixing}
\end{figure}

Nevertheless, at least where the uncertainty in signal measurements 
is low, we can predict the orientation from the scattering scattering 
by utilising computational techniques such as neural networks. We 
thus utilised the Python machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a neural network for identifying the 
dimer's orientation from its light scattering signal. The network 
was trained by generating a database of random orientation vectors, 
calculating the corresponding light scattering signals, and then 
using the network to estimate the probability of a given signal coming 
from a dimer in a given reference orientation. The network's loss 
function was evaluated and used to improve the estimation, the 
network being trained until the improvement in the loss function 
was less than 0.0001.

@@

The relative entropy is a measure if how spread out our distribution
is. With larger values of $D_{KL}$ indicating that our prediction is 
spread out so the neural net is unsure about $\hat{\bf n}_{best}$. 
Whereas a low value of $D_{KL}$ means that our model is more confident 
in $\hat{\bf n}_{best}$. The only draw back being that the relative 
entropy does nothing to indicate accuracy, only how closely our 
distribution matches our ideal result. This will be better 
exemplified later in \ref{sec:test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Brownian Simulation}
\label{sec:brownian}

@@

\subsection{Choice on the number of detectors}
When all of the detectors lie in the same plane the expected signal 
can appear identical despite the dimer being in completely different orientations. It should be noted that these pairs are reflected in 
one or more axis which suggests that these are due to the arrangement 
of our detectors. More specifically, if the detectors are placed say 
in the x-y plane then only when the dimer is pointed nearly fully 
upright will the expected signal be entirely unique. This is not an 
issue with typical analytical static light scattering experiments 
because the signal being collected is averaged over a population of 
particles all in different orientations. Whereas in our set up we are concerned with the scattering of a single particle whose scattering is orientationally dependant.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Testing the Neural Network}
\label{sec:test}
We applied Eq.~\eqref{eq:bayes}, taking the reference orientation 
with the highest probability  as our estimate of the dimer's 
instantaneous orientation $\hat{\bf n}_{est}$. To visualise the 
model's performance we plotted the orientation components of our 
estimation $\hat{\bf n}_{\alpha}$ and the dimer's \emph{actual} 
instantaneous orientation $\hat{\bf s}$ versus time. For comparison, 
we also plotted the orientation components of the closest reference orientation, $\hat{\bf n}_{best}$. Assuming a uniform prior of the 
reference orientations $p(\hat{\bf n}_\alpha)$ the neural network's predictions ($\hat{\bf n}_{\alpha}$ from Eq.~\eqref{eq:bayes}) are 
at times reasonable, but there are significant large and random jumps 
away from the correct result (Fig.~\ref{fig:uniform}). 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{fig5.png}
	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity. Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the blue line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$. Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
	\label{fig:uniform}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Accounting for sources of error in light scattering measurements}
When it comes to analysing light scattering from any size particle,
error analysis becomes a significant factor. Typically this can be 
accounted for by averaging over long periods of time to get an 
assessment of the steady state conditions of the target particle. 
However in our case where we wish to know the instantaneous 
orientation, we instead have to rely on our understanding of how 
uncertainty can effect our model's performance. We identified two 
areas which are likely sources of error in our estimation: firstly, 
an incorrect modelling of the target particle, and secondly, signal 
noise arising from experimental factors. We highlight how we address 
these areas below. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Impact of incorrect dimer sizing}
\label{sec:lam}

@@

Notably the increasing the number of reference orientations had a greater effect when our neural network was trained on a 1:1.95 dimer than a 1:2.05 dimer. This suggests that overshooting our size estimate will be less detrimental to our estimation. Notably if the our sizing is off the neural network does not predict a smooth motion within the trap; instead predicting that the dimer is jumping back and forth between different orientations. This suggest that we can narrow down our estimate of the particle's size by assessing how the dimer is reorienting within the trap, as we should expect a smooth continuos prediction. Since we are working with a spherical dimer it also stands to reason that techniques such as image analysis could be used in part to address this, so long as the trapped entity is sufficiently illuminated. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Impact of measurement noise on model predictions}
\label{sec:epsilon}

@@

Conventional calibration methods will not be adequate however, while
angular motion can be detected using complicated set ups \cite{Bang2020},
this can only provide an estimate of the magnitude of a particle's 
angular motion. This is perfectly fine in cases where the rotational
motion is stochastic and intuitively predictable (i.e. a dimer in a
vertical orientation undergoing stochastic Brownian motion), but in 
the case the the motion is instead periodic characterising it requires
instantaneous measurements of the particle's motion.
